---
title: "Linear Regression analysis of estimating age of giving birth by using marriage age"
author: "Shu Liu, Xinrui Wang, Github: https://github.com/Shu2000/STA304"
date: "Oct. 19th"
output:
  pdf_document: 
      latex_engine: xelatex
  html_document:
    df_print: paged
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#### Workspace set-up ####
library(janitor)
library(tidyverse)

# Load the data dictionary and the raw data and correct the variable names
raw_data <- read_csv("AAWY.csv")
dict <- read_lines("gss_dict.txt", skip = 18) # skip is because of preamble content
# Now we need the labels because these are the actual responses that we need
labels_raw <- read_file("gss_labels.txt")


#### Set-up the dictionary ####
# What we want is a variable name and a variable definition
variable_descriptions <- as_tibble(dict) %>% 
  filter(value!="}") %>% 
  mutate(value = str_replace(value, ".+%[0-9].*f[ ]{2,}", "")) %>% 
  mutate(value = str_remove_all(value, "\"")) %>% 
  rename(variable_description = value) %>% 
  bind_cols(tibble(variable_name = colnames(raw_data)[-1]))
 
# Now we want a variable name and the possible values
labels_raw_tibble <- as_tibble(str_split(labels_raw, ";")[[1]]) %>% 
  filter(row_number()!=1) %>% 
  mutate(value = str_remove(value, "\nlabel define ")) %>% 
  mutate(value = str_replace(value, "[ ]{2,}", "XXX")) %>% 
  mutate(splits = str_split(value, "XXX")) %>% 
  rowwise() %>% 
  mutate(variable_name = splits[1], cases = splits[2]) %>% 
  mutate(cases = str_replace_all(cases, "\n [ ]{2,}", "")) %>%
  select(variable_name, cases) %>% 
  drop_na()

# Now we have the variable name and the different options e.g. age and 0-9, 10-19, etc.
labels_raw_tibble <- labels_raw_tibble %>% 
  mutate(splits = str_split(cases, "[ ]{0,}\"[ ]{0,}"))

# The function sets up the regex (I know, I know, but eh: https://xkcd.com/208/)
add_cw_text <- function(x, y){
  if(!is.na(as.numeric(x))){
    x_new <- paste0(y, "==", x,"~")
  }
  else{
    x_new <- paste0("\"",x,"\",")
  }
  return(x_new)
}

# The function will be in the row, but it'll get the job done
cw_statements <- labels_raw_tibble %>% 
  rowwise() %>% 
  mutate(splits_with_cw_text = list(modify(splits, add_cw_text, y = variable_name))) %>% 
  mutate(cw_statement = paste(splits_with_cw_text, collapse = "")) %>% 
  mutate(cw_statement = paste0("case_when(", cw_statement,"TRUE~\"NA\")")) %>% 
  mutate(cw_statement = str_replace(cw_statement, ",\"\",",",")) %>% 
  select(variable_name, cw_statement)
# So for every variable we now have a case_when() statement that will convert 
# from the number to the actual response.

# Just do some finally cleanup of the regex.
cw_statements <- 
  cw_statements %>% 
  mutate(variable_name = str_remove_all(variable_name, "\\r")) %>% 
  mutate(cw_statement = str_remove_all(cw_statement, "\\r"))


#### Apply that dictionary to the raw data ####
# Pull out a bunch of variables and then apply the case when statement for the categorical variables
gss <- raw_data %>% 
  select(CASEID, 
         agedc, 
         achd_1c, 
         achdmpl, 
         totchdc, 
         acu0c,
         agema1c,
         achb1c,
         rsh_131a,
         arretwk,
         slm_01, 
         sex, 
         brthcan, 
         brthfcan,
         brthmcan,
         brthmacr,
         brthprvc,
         yrarri,
         prv, 
         region, 
         luc_rst, 
         marstat, 
         amb_01, 
         vismin, 
         alndimmg,
         bpr_16, 
         bpr_19,
         ehg3_01b, 
         odr_10, 
         livarr12, 
         dwelc, 
         hsdsizec,
         brthpcan,
         brtpprvc, 
         visminpr,
         rsh_125a, 
         eop_200,
         uhw_16gr,
         lmam_01, 
         acmpryr,
         srh_110,
         srh_115,
         religflg, 
         rlr_110,
         lanhome, 
         lan_01,
         famincg2, 
         ttlincg2, 
         noc1610, 
         cc_20_1,
         cc_30_1,
         ccmoc1c,
         cor_031,
         cor_041,
         cu0rnkc,
         pr_cl,
         chh0014c,
         nochricc,
         grndpa,
         gparliv,
         evermar,
         ma0_220,
         nmarevrc,
         ree_02,
         rsh_131b,
         rto_101,
         rto_110,
         rto_120,
         rtw_300,
         sts_410,
         csp_105,
         csp_110a,
         csp_110b,
         csp_110c,
         csp_110d,
         csp_160,
         fi_110) %>% 
  mutate_at(vars(agedc:fi_110), .funs = funs(ifelse(.>=96, NA, .))) %>% 
  mutate_at(.vars = vars(sex:fi_110),
            .funs = funs(eval(parse(text = cw_statements %>%
                                      filter(variable_name==deparse(substitute(.))) %>%
                                      select(cw_statement) %>%
                                      pull()))))

# Fix the names
gss <- gss %>% 
  clean_names() %>% 
  rename(age = agedc,
         age_first_child = achd_1c,
         age_youngest_child_under_6 = achdmpl,
         total_children = totchdc,
         age_start_relationship = acu0c,
         age_at_first_marriage = agema1c,
         age_at_first_birth = achb1c,
         distance_between_houses = rsh_131a,
         age_youngest_child_returned_work = arretwk,
         feelings_life = slm_01,
         sex = sex,
         place_birth_canada = brthcan,
         place_birth_father = brthfcan,
         place_birth_mother = brthmcan,
         place_birth_macro_region = brthmacr,
         place_birth_province = brthprvc,
         year_arrived_canada = yrarri,
         province = prv,
         region = region,
         pop_center = luc_rst,
         marital_status = marstat,
         aboriginal = amb_01,
         vis_minority = vismin,
         age_immigration = alndimmg,
         landed_immigrant = bpr_16,
         citizenship_status = bpr_19,
         education = ehg3_01b,
         own_rent = odr_10,
         living_arrangement = livarr12,
         hh_type = dwelc,
         hh_size = hsdsizec,
         partner_birth_country = brthpcan,
         partner_birth_province = brtpprvc,
         partner_vis_minority = visminpr,
         partner_sex = rsh_125a,
         partner_education = eop_200,
         average_hours_worked = uhw_16gr,
         worked_last_week = lmam_01,
         partner_main_activity = acmpryr,
         self_rated_health = srh_110,
         self_rated_mental_health = srh_115,
         religion_has_affiliation = religflg,
         regilion_importance = rlr_110,
         language_home = lanhome,
         language_knowledge = lan_01,
         income_family = famincg2,
         income_respondent = ttlincg2,
         occupation = noc1610,
         childcare_regular = cc_20_1,
         childcare_type = cc_30_1,
         childcare_monthly_cost = ccmoc1c,
         ever_fathered_child = cor_031,
         ever_given_birth = cor_041,
         number_of_current_union = cu0rnkc,
         lives_with_partner = pr_cl,
         children_in_household = chh0014c,
         number_total_children_intention = nochricc,
         has_grandchildren = grndpa,
         grandparents_still_living = gparliv,
         ever_married = evermar,
         current_marriage_is_first = ma0_220,
         number_marriages = nmarevrc,
         religion_participation = ree_02,
         partner_location_residence = rsh_131b,
         full_part_time_work = rto_101,
         time_off_work_birth = rto_110,
         reason_no_time_off_birth = rto_120,
         returned_same_job = rtw_300,
         satisfied_time_children = sts_410,
         provide_or_receive_fin_supp = csp_105,
         fin_supp_child_supp = csp_110a,
         fin_supp_child_exp = csp_110b,
         fin_supp_lump = csp_110c,
         fin_supp_other = csp_110d,
         fin_supp_agreement = csp_160,
         future_children_intention = fi_110) 

#### Clean up ####
gss <- gss %>% 
  mutate_at(vars(age:future_children_intention), 
            .funs = funs(ifelse(.=="Valid skip"|.=="Refusal"|.=="Not stated", "NA", .))) 

gss <- gss %>% 
  mutate(is_male = ifelse(sex=="Male", 1, 0)) 

gss <- gss %>% 
  mutate_at(vars(fin_supp_child_supp:fin_supp_other), .funs = funs(case_when(
    .=="Yes"~1,
    .=="No"~0,
    .=="NA"~as.numeric(NA)
  )))

main_act <- raw_data %>% 
  mutate(main_activity = case_when(
    mpl_105a=="Yes"~ "Working at a paid job/business",
    mpl_105b=="Yes" ~ "Looking for paid work",
    mpl_105c=="Yes" ~ "Going to school",
    mpl_105d=="Yes" ~ "Caring for children",
    mpl_105e=="Yes" ~ "Household work", 
    mpl_105i=="Yes" ~ "Other", 
    TRUE~ "NA")) %>% 
  select(main_activity) %>% 
  pull()

age_diff <- raw_data %>% 
  select(marstat, aprcu0c, adfgrma0) %>% 
  mutate_at(.vars = vars(aprcu0c:adfgrma0),
            .funs = funs(eval(parse(text = cw_statements %>%
                                      filter(variable_name==deparse(substitute(.))) %>%
                                      select(cw_statement) %>%
                                      pull())))) %>% 
  mutate(age_diff = ifelse(marstat=="Living common-law", aprcu0c, adfgrma0)) %>% 
  mutate_at(vars(age_diff), .funs = funs(ifelse(.=="Valid skip"|.=="Refusal"|.=="Not stated", "NA", .))) %>% 
  select(age_diff) %>% 
  pull()

gss <- gss %>% mutate(main_activity = main_act, age_diff = age_diff)

# Change some from strings into numbers
gss <- gss %>% 
  rowwise() %>% 
  mutate(hh_size = str_remove(string = hh_size, pattern = "\\ .*")) %>% 
  mutate(hh_size = case_when(
    hh_size=="One" ~ 1,
    hh_size=="Two" ~ 2,
    hh_size=="Three" ~ 3,
    hh_size=="Four" ~ 4,
    hh_size=="Five" ~ 5,
    hh_size=="Six" ~ 6
  )) 

gss <- gss %>% 
  rowwise() %>% 
  mutate(number_marriages = str_remove(string = number_marriages, pattern = "\\ .*")) %>% 
  mutate(number_marriages = case_when(
    number_marriages=="No" ~ 0,
    number_marriages=="One" ~ 1,
    number_marriages=="Two" ~ 2,
    number_marriages=="Three" ~ 3,
    number_marriages=="Four" ~ 4
  )) 

gss <- gss %>% 
  rowwise() %>% 
  mutate(number_total_children_known = ifelse(number_total_children_intention=="Don't know"|number_total_children_intention=="NA", 0, 1)) %>% 
  mutate(number_total_children_intention = str_remove(string = number_total_children_intention, pattern = "\\ .*")) %>% 
  mutate(number_total_children_intention = case_when(
    number_total_children_intention=="None" ~ 0,
    number_total_children_intention=="One" ~ 1,
    number_total_children_intention=="Two" ~ 2,
    number_total_children_intention=="Three" ~ 3,
    number_total_children_intention=="Four" ~ 4,
    number_total_children_intention=="Don't" ~ as.numeric(NA)
  )) 

write_csv(gss, "gss.csv")

```


## Abstract

This study aimed to analyze the relationship between marriage and the ages of getting the first child. We have developed a linear regression model on marriage age to get the first child from the gss 2017 data set. We find that marriage is positively correlated to the age of getting the first child with strong statistical evidence. Together with some analysis of using observational data as the base of modelling, we have also suggested a better model that can be used in future studies to show a more accurate result.  

## Introduction

The relationship between marriage and giving birth is a frequently discussed topic in news or reports. A common sense is that the younger people get married, the earlier they will get the first birth. However, some state that when younger people get married, they prefer not to get pregnant which may restrict their works or lives, and thus forming a panic of getting pregnant which will be a long lasting influences that may stimulate them to have the first birth older than the others. 

Consequently, the goal of this analysis is to discover the relationship between the age at marriage and the age at first birth as well as the impact of marriage age imposed on reproduction age. 

A supervised study will be carried out between these two variables in this case. The model may help estimating the birth rate or population in the future by collecting marriage data. Discussions of limitations in the analysis will also be included. 

## Data

The data used in the analysis is from Canadian General Social Survey(GSS) at a particular year 2017. The reason of choosing the data is the large sample size survey conducted by GSS, the larger the sample, the better estimation of the population will be. 

The data is downloaded on the Chass Website. To use the data, a procedure of cleaning data is processed previously, a data frame is produced finally. In this case, the overall society is the target population for this analysis and the surveyed people in this dataset are the sample to estimate the population.

As this analysis is going to find how reproduction age is affected by marriage age, the response variable is then the age of first giving birth. 

In the GSS dataset, another variable can be also used for estimating reproduction age, age of first child. This can also reflect the relationship between marriage and giving birth. However, in this case, the age at first birth shows the relation in a more direct way than the age of first child. Although age of first child can be also used in the model, the analysis will still not use this variable. 

As a result, to achieve the goal, two variables: age_at_first_marriage and age_at_first_birth will be focused.

## Model

Since the focused two variables are all numerical variables, a scatterplot will be plotted to firstly discover their relationship. To make an estimation of the overall population, a simple linear regression will be made through least square method by rcode, which can help showing that whether they relate to each other linearly or not. This will result in a model, following the formula: 

                            y=b0+b1∗x
                            
The analysis is aiming at discovering the influences of marriage age imposed on age of giving birth. Hence, y is the response variable, the age of giving first birth, x is the explanatory variable, the age of first marriage. b0 stands for the intercept and b1 represents the rate of change. The resulted model can give an overall estimation through these two variables. 

P-value testing will also be analyzed in the following sections to check if the model can fit the data. 

The overall model will be generated through R code thoroughly, with additional packages for helping. Figures and tables will then be produced. 

Despite SLR, other models such as multiple linear regression can be also used. That's because reproduction age may be influenced by mutiple factors, and thus more explanatory variables can be considered. That's the limitations of SLR. 

A diagnostic issue should be analyzed through analyzing cook distance. In this case, Di = 4/n = 4/20600, the data with a cook distance less than 4/20600 will be selected in the model to reduce influence of outliers. In addition, it can be noticed that there are influential points, where age at first marriage greater than 40 but has a low age of giving birth. To eliminate these outliers, the dataset should be further cleaned. 


## Results
```{r,message=FALSE, warning=FALSE,echo=FALSE}
library(ggplot2)
library(dplyr)

attach(gss)

lmbr=lm(age_at_first_birth~age_at_first_marriage,gss)

dg <- cbind(age_at_first_marriage,age_at_first_birth, lmbr$residuals,rstandard(lmbr),hatvalues(lmbr),cooks.distance(lmbr))
colnames(dg)=c("age_at_first_marriage","age_at_first_birth","Residuals", "Std Resids", "Leverage","Cook")

Di = 4/20600

diag <- data.frame(dg)

Gss <- filter(diag, Cook < Di)

GSS <- filter(Gss, age_at_first_marriage < 40)

ggplot(GSS, aes(x=age_at_first_marriage,y=age_at_first_birth))+geom_point()+geom_smooth(method = 'lm', se = FALSE)+labs(title = 'Figure1')+theme(plot.title=element_text(hjust = 0.5))

lm1=lm(age_at_first_birth~age_at_first_marriage,GSS)

summary(lm1)

anova(lm1)

confint(lm1, level=0.95)



```

In figure 1, a scatterplot with a linear regression line is plotted. Two variables, age at first birth and age at first marriage have a positive relationship, which means when marriage age rises, the reproduction age will also increase. 

## Discussion
This analysis aimed to discover a relationship between the age of marriage(age_first_marriage) and the age of getting a first child(age_at_first_birth), so the result above will be break into two parts for discussion. First, to discuss the statistical significance of the slope coefficient and second, examine the model's goodness of fit.

The null hypothesis in the analysis is that the age of first marriage does not affect the age of getting the first child. The alternative hypothesis is that the age of first marriage affects the age of the first child. The two variables should have a positive relationship. The result showed in the previous section gives us a positive value of the slope coefficient, which is 0.82737. This slope coefficient shows a ten years increase in age of first marriage is associated with a 8.27 years increase in the age of getting the first child, on average.

The t-statistics of the slope coefficient is 50.32, far more extensive than the 95% rejection critical value of 1.96. Thus, we have sufficient evidence to reject the null hypothesis and infer that the age of first marriage affects the age of getting the first child. 

However, when we look at the model's fitness to the data, the R^2 value is only 0.441, which means that variations of ages of first marriage can explain only 44% of the first child's variations in ages. The low R^2 value with a high statistical significance shows that there might be some possible omitted variables, which we will discuss in sections below.

# Weaknesses
This study's major weakness is that the data that we have used to model is observational data rather than experimental data. Using observational data to infer causal relationships is an issue since the relationship we have identified can also be affected by unobserved variables. Unobserved variables could also explain why even the t-statistics is quite significant, and we still got a low level of fitness. 

For example, family income can be an essential factor that affects both ages of getting first child and age of marriage. People are more likely to get married earlier when they do not have much income, and they are more likely to have a child more before. Another example can be the educational level. When people are more educated, they are less likely to get married early and get a child first.  These examples provide insights into how an unobserved variable could affect both dependent and independent variables we are interested in studying and affect the overall result. 

Hence, due to those possible unobserved variables, the model's effect may not be the true effect of the age of first marriage on the age of getting the first child. 

# Next Steps
We have identified a significant weakness in our analysis in the weakness section because this analysis is suffering from endogeneity bias. The best way to remove the bias is to randomly assign treated units with different ages of getting the first child, which is not realistic. Therefore, as an alternative,  to further investigate the true relationship between the variables to better our goals, our next step is to develop a multiple linear regression model that controls the omitted variables such as educational level, income level, ethnicity, and more. By controlling those omitted variables, we can solve some of the problems by using observational data. It is also impossible to include all variables into the model; possible reasons such as pregnancy-related disease are challenging to measure and included in the analysis. However, by introducing those variables, we are confident that the new model will provide us with a more accurate relationship between the age of getting the first child and the age of marriage. 


## References

1. Sam Firke (2020). janitor: Simple Tools for Examining and
  Cleaning Dirty Data. R package version 2.0.1.
  https://CRAN.R-project.org/package=janitor
  
2. H. Wickham. ggplot2: Elegant Graphics for Data Analysis.
  Springer-Verlag New York, 2016.

3. Hadley Wickham, Romain Fran?ois, Lionel Henry and Kirill
  Müller (2020). dplyr: A Grammar of Data Manipulation. R
  package version 1.0.2.